---
title: "Eating Out"
author: "Jessica Villa"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The data comes from the 1998 General Social Survey answer this question

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Methods

The response variable is categorical. Which is, did you dine out last week? The explanatory variable is race: another categorical variable. The reponse variavle fits under yes or no thus, it's categorical. For race, we know from past experience that it is for sure a categorical variable. So we use cat~cat. Thus, we'll a chi squared test and a fisher exact test. The inferential test will be done by the chi squared test & the fisher exact test. 


# Descripitve summaries 

## Graphical summary 

### Barchart

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")
``` 

White tends to be the outlier in this bar plot. Since their percent of dining out is the highest (close to 85% of white people dine out according to this bar plot). Black & other are relatively the same with 'other' only being a slightly higher. There is a clear difference so far since one of the 3 examples is way higher than the other 2. 

## Numerical summary 

```{r}
table1 <- xtabs(~race + eatout,data=mydata)
rowPerc(table1)
colPerc(table1)
```
This table shows that there is not much of a difference between black race 69.04 and other races 72.12. We do see a slight significant difference with the white race 83.11, so when we compare 69.04 to 83.11 we see that there is an increase in white race eating out compared to the black and other races.  

# Results 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi-squared test shows that there is a significant difference of the white race eating out compared to the other races.The p-value is 0.000002287. The p-value is way too small so we have to reject the null hypothesis. This shows that there is a difference between race and how frequent they dine out. 

```{r}
fisherXY(mydata$race,mydata$eatout)
```
There is a difference in frequency of the races. The p-value is low 0.0721. The other p-value is 1.183e-05. Since the p-value is less than 0.00721 we fail to reject the null hypothesis.   

# Discussion and Conclusion


We can conclude that white people are more likely to eat out than black and other races. The reason we can conclude this is because we looked at and compared the charts, tables, and the fisher test. When we compared these 3 charts they concluded the same information which was that the white race had a higher percent then the other races. Both tests have a p-value less then 0.00721 and therefore we fail to reject the null hypothesis.   

