---
title: "Eating Out"
author: "Jessica Villa"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The data comes from the 1998 General Social Survey answer this question

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Methods

Response variable is categorical: Dine out last week? Yes or no?

Explanatory variable is race, it, too, is a categorical variable. 
So we use cat~cat 

Inference: chi-squared and fisher exact tests 

# Descripitve summaries 

## Graphical summary 

### Barchart

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```

By examining the chart there is not much of a difference between black and other race the percent is almost the same. We do see a difference with white race their percent is less then the others. Meaning that the white race eats out more then the other races. 

## Numerical summary 

```{r}
table1 <- xtabs(~race + eatout,data=mydata)
rowPerc(table1)
colPerc(table1)
```
This table shows that there is not much of a difference between black race 69.04 and other races 72.12. We do see a slight significant difference with the white race 83.11, so when we compare 69.04 to 83.11 we see that there is an increase in white race eating out compared to the black and other races.  

# Results 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Chi-squared test shows that 


