---
title: "Eating Out"
author: "Jessica Villa"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The data comes from the 1998 General Social Survey answer this question

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

# Methods

The response variable is categorical. Which is, did you dine out last week? The explanatory variable is race: another categorical variable. The reponse variavle fits under yes or no thus, it's categorical. For race, we know from past experience that it is for sure a categorical variable. So we use cat~cat. Thus, we'll a chi squared test and a fisher exact test. The inferential test will be done by the chi squared test & the fisher exact test. 


# Descripitve summaries 

## Graphical summary 

### Barchart

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")
``` 

White tends to be the outlier in this bar plot. Since their percent of dining out is the highest (close to 85% of white people dine out according to this bar plot). Black & other are relatively the same with 'other' only being a slightly higher. There is a clear difference so far since one of the 3 examples is way higher than the other 2. 

## Numerical summary 

````{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```
Based off of the data table on top, more white people dine out compared to the other two (black & other). White's dine out more by 14.07 compared to black people. White's dine out more by 10.99 compared to 'other' people. This data table only further proves that there is a difference between the 3 options. White dines out more. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is  0.000002287. The p-value is too small so we have to reject the null hypothesis. The p-value for the Fisher's Exact test is .000004633. This only further supports my choice to reject the null hypothesis. There IS a difference between race and how frequent they dine out. That's what the chi squared test gave us. 

```{r}
fisherXY(mydata$race,mydata$eatout)
```

There is a statistical difference between at least two races in regard to the frequency of eating out. I reject the null hypothesis since the p-value is 0.000002287, much less than 0.05.  When comparing white to black individuals,  whites tend to eat out more frequently with a ratio of  2.2:1 when compared to other individuals. Whites are 1.9 times more likely to eat out. By using the fisher exact test, I am able to show that whites are more likely to eat out compared to blacks and others. Therefore, with certainty, I am able to reject the null hypothesis. Since the fisher exact test didn't print out anything for black & other.They have no statistical difference. That's the difference: White is the clear outlier. Black and other aren't different enough to have a statistical result of their own. While the white race clearly has different results compared to black & other. 
  

# Discussion and Conclusion


We can conclude that white people are more likely to eat out than black and other races. The reason we can conclude this is because we looked at and compared the charts, tables, and the fisher test. When we compared these 3 charts they concluded the same information which was that the white race had a higher percent then the other races. Both tests have a p-value less then 0.00721 and therefore we fail to reject the null hypothesis.   

